On Etiology of Cybersecurity

### 2020-06-25 reports/paper_0/main.tex
implement Luca's feedback

\newtheorem{definition}{Definition}%[section]
\newtheorem{hypothesis}{Hypothesis}%[section]

@@

\section{Introduction}\label{sec:intro}
In \autocite{Herley2016unfalsifiability}, Cormac Herley explores what he calls
``an asymmetry in computer security'', which he defines as follows: ``Things
can be declared insecure by observation, but not the reverse. There is no
observation that allows us to declare an arbitrary system or technique
secure''. Herley then uses this argument to show that ``claims that any measure
is necessary for security are empirically unfalsifiable''. Given that, any
theory which is not falsifiable by an empirical experiment is well
known\footnote{``A theory which is not refutable by any conceivable event is
nonscientific. Irrefutability is not a virtue of a theory (as people often
think) but a vice.'' -- Karl Popper, Conjectures and
Refutations \autocite{popper1962conjectures}} to be nonscientific (i.e.
unfalsifiability is a fallacy of a theory), Herley concludes that there is no
scientific theory on cybersecurity; which means that cybersecurity lays in the
realm of pseudo-sciences \autocite{Herley2016usenixvideo}.  Herley, e.g.
in \autocite{Herley2017justifying}, discusses the implications of a
nonscientific approach to cybersecurity, and highlights the tremendous impact on
all the scientific research and engineering of systems; leading often to
terrorism and wars, and wasting of resources in useless protections or
overspending.  While the criticism is investigated
in \autocite{Herley2016unfalsifiability}, no solution is provided nor envisioned.  On the
contrary, the goal of this work is to lay the foundations of a
scientific cybersecurity theory.

We consider the problem raised by Herley not confined to
``computer security'' but to any abstract system (so that our theory may hold
for any sound implementation such as networks, mechanical, cyber, or
cyber-physical system, or even a single computer or a single device such as an
hard-drive).  There is also an apparent inconsistency in
\autocite{Herley2016unfalsifiability} that we seek to clarify before following (as
we agree) the scientific path draw by Herley: cybersecurity is defined as an
abstract property in many formal approaches to the investigation of the
security of systems, and the security of the design of a formally verified
protocol is indeed falsifiable (against the security properties verified).  For
example, in the protocol verification community, security is often defined as a
formalization of the high-level properties confidentiality, integrity, and
availability. The problem in such approaches is not the definition of what
cybersecurity is but the generality of the results, since they tackle a specific
step (and not the first) of the engineering process (of systems).  Therefore,
the theories underlying the verification are based on assumptions which

@@

We are now in the position to define what a secure system is (with respect to
the $\abf$-theory) and, based on that definition, what the security risk is and
how to quantify it in a risk matrix.
\begin{definition}{\bf Security of a System or an Agent --}\label{def:security}
	A secure system is a system where SE-$1$, SE-$2$, and SE-$3$ holds for
	each agent composing the system.
\end{definition}
The ISO 31000 consider risk as the ``effect of uncertainty on objectives'' and
refers both to positive and negative consequences of uncertainty.
Accordingly, we consider risk as follows. 
\begin{definition}{\bf Risk --}
The whole space of potential designs of a specification with respect to the
	$\abf$-theory.
\end{definition}
The definition of Risk leads to the risk matrix in
Figure~\ref{fig:quantitative}, defined as follows.
\begin{definition}{\bf Risk Matrix --}
	The risk matrix is a function between the three relations
	$s=\langle\rcc(\factRegion,\beliefRegion),\rcc(\factRegion,\assertionRegion),\rcc(\beliefRegion,\assertionRegion)\rangle$,
	where the maximum risk is defined by the $DR$ relation between the
	three groups of Regions, and the minimum risk by the $EQ$ relation over
	the same Regions. In between the two extremes, the granularity of
	possible intermediate configuration is defined by the calculus used
	(RCC5 in our case).
\end{definition}
The risk matrix is often defined as a function between 
likelihood and impact of attacks. We suggest
that security weaknesses are all equally likely to be exploited
if there's a connection between the weakness and the asset that
an attacker wants to impact. Therefore, a risk matrix should
capture the number of insecure configurations of a system, rather
than predicating over likelihoods of weaknesses exploitation.

@@

\paragraph{Port Weaknesses}
From our definition of ports, it follows that there exist only the following \emph{six}
types of weaknesses, generating six types of insecure port in RCC5 (the
notation is reported for input-ports, i.e. on the lef-hand side of the arrow):
\begin{enumerate}[start=1, label={W\arabic*)}]
	\item \emph{replace port} ($s\dropport a$), where assertions reaches
		the port but do not pass the boundary of the agent (i.e. do not
		become belief of the agent)
	\item \emph{drop port} ($s\dropport a$), where assertions reaches the
		port but do not pass the boundary of the agent (i.e. do not
		become belief of the agent)
	\item \emph{insertion port} ($s\insertport a$), where some new
		information is believed by $a$ as incoming from the port but
		$s$ didn't send it
	\item \emph{injection port} ($s\injectionport a$), where information
		coming from $s$ is substituted with new information which
		becomes believed by $a$
	\item \emph{selective port}, where some information passes the port and
		part is either:
	\begin{enumerate}[start=1, label={W4.\arabic*)}]
		\item drop ($s\selectivedropport a$), 
		\item drop+insert ($s\selectiveinsertport a$).
	\end{enumerate}
\end{enumerate}

@@

\begin{enumerate}[start=6, label={W\arabic*)}]
	\item \emph{secure output port} and \emph{input drop port} ($s\sdcom r$), 
	\item \emph{secure output port} and \emph{input insertion port} ($s\sicom r$), 
	\item \emph{output drop port} and \emph{input drop port} ($s\ddcom r$)
	\item \emph{output drop port} and \emph{input insertion port} ($s\dicom r$)
	\item \emph{output drop port} and \emph{input secure port} ($s\dscom r$)
	\item \emph{output injection port} and \emph{input secure port} ($s\jscom r$)
\end{enumerate}

@@

\section*{Acknowledgment}
We thank Katia Santac\`a for the help in the development of the initial ideas, and for
the helpful discussions that allowed us to make this step.

>>>>

\newtheorem{definition}{Definition}%[section]
\newtheorem{hypothesis}{Hypothesis}%[section]
\newtheorem{theorem}{Theorem}%[section]

@@

\section{Introduction}\label{sec:intro}
A scientific theory is an explanation of a phenomenon such that the explanation
follows the scientific method. The scientific method is an \emph{empirical}
method that aims at mitigating potential fallacies in theories.  K. Popper
famously argued (e.g. in \autocite{Popper1959logic}) that a scientific theory
can never been verified but only falsified, that a theory should not be
conceived by using the principle of induction\footnote{A.  Einstein wrote to K.
Popper: ``[...] and I think (like you, by the way) that theory cannot be
fabricated out of the results of observation, but that it can only be
invented.''\autocite{Popper1959logic}}, and that empirical experiments should
be considered as the only evidence to support the non-falseness of a scientific
theory. If we consider a natural (real-world) phenomenon, such as gravity, we
can think of a scientific theory as an \emph{abstract} mathematical formula (or
a set of formulas) that axiomatically define a certain (natural) phenomenon in
a specific algebraic structure. While such a theory can be proved to be correct
within the mathematical setting it has been conceived in (e.g. with respect to
other axioms of the algebraic structure where it has been postulated, or with
respect to other background theories), no proof (for soundness or completeness)
can be carried out to verify that the theory correctly abstracts  the
phenomenon it describes. Similarly, the results of the application of a
scientific theory (i.e. its predictions) can be verified in the abstract
mathematical structure but how can we be sure that the theory correctly
characterize the phenomenon it describes? With his famous ``demarcation
principle'', K. Popper draw a line between scientific and pseudo-scientific
theories by proposing that \emph{any} scientific theory can be considered
correct as long as no empirical experiment falsify it. For example, the theory
of gravity, in its mathematical setting, can predict how object attracts with
each other, and those predictions can be tested by empirical experiment. If the
theory correctly abstracts the phenomenon no empirical experiment should
falsify the predictions of that theory.

In \autocite{Herley2016unfalsifiability}, Cormac Herley explores what he calls
``an asymmetry in computer security'', which he defines as follows: ``Things
can be declared insecure by observation, but not the reverse. There is no
observation that allows us to declare an arbitrary system or technique
secure''. His intuition is that there is no scientific theory that can predict
the security of a system, nor a theory that can predict all possible
insecurity of a system (which, by negation it would become a theory on security). 
Herley then uses this argument to show that ``claims that any measure
is necessary for security are empirically unfalsifiable''. Given that, any
theory which is not falsifiable by an empirical experiment is well
known\footnote{``A theory which is not refutable by any conceivable event is
nonscientific. Irrefutability is not a virtue of a theory (as people often
think) but a vice.'' -- Karl Popper, Conjectures and
Refutations \autocite{popper1962conjectures}} to be nonscientific (i.e.
unfalsifiability is a fallacy of a theory), Herley concludes that there is no
scientific theory on cybersecurity; which means that cybersecurity lays in the
realm of pseudo-sciences \autocite{Herley2016usenixvideo}.  Herley, e.g.
in \autocite{Herley2017justifying}, discusses the implications of a
nonscientific approach to cybersecurity, and highlights the tremendous impact on
all the scientific research and engineering of systems; leading often to
terrorism and wars, and wasting of resources in useless protections or
overspending.  While the criticism is investigated
in \autocite{Herley2016unfalsifiability}, no solution is provided nor envisioned.  On the
contrary, the goal of this work is to lay the foundations of a
scientific cybersecurity theory.

We consider the problem raised by Herley not confined to ``computer security''
but to any abstract system (so that our theory may hold for any sound
implementation such as networks, mechanical, cyber, or cyber-physical system,
or even a single computer or a single device such as an hard-drive).  There is
also an apparent inconsistency in \autocite{Herley2016unfalsifiability} that we
seek to clarify before following (as we agree) the scientific path draw by
Herley: cybersecurity is defined as an abstract property in many formal
approaches to the investigation of the security of systems, and the security of
the design of a formally verified protocol is indeed falsifiable (against the
security properties verified).  For example, in the protocol verification
community, security is often defined as a formalization of the high-level
properties confidentiality, integrity, and authentication. The problem in such
approaches is not the definition of what cybersecurity is but the generality of
the results, since they tackle a specific step\footnote{In fact, many standards
(e.g. ) used for the engineering of systems starts with the requirements
definition, and physical and functional architecture design, before the logic
of protocols is even considered.} of the engineering process (of systems)
without considering many aspects related to the implementation of a protocol
that may generate security vulnerabilities later in the development process. An
infamous example is the KRAK attack on the WPA2 4-way handshake protocol where
the authors wrote\autocite{krak2020web}: ``our attacks do not violate the security properties proven
in formal analysis of the 4-way handshake'' but ``the problem is that the
proofs do not model key installation.''.
The theories underlying the verification of security protocols are based on assumptions which

@@

We are now in the position to define what a secure system is (with respect to
the $\abf$-theory) and, based on that definition, what the security risk is and
how to quantify it in a risk matrix (where a risk matrix is usually defined as
a function between the likelihood and the impact of all potential attacks).
\begin{definition}{\bf Security of a System or an Agent --}\label{def:security}
	A secure system is a system where SE-$1$, SE-$2$, and SE-$3$ holds for
	each agent composing the system.
\end{definition}
The ISO 31000 consider risk as the ``effect of uncertainty on objectives'' and
refers both to positive and negative consequences of uncertainty.
Accordingly, we consider risk as follows. 
\begin{definition}{\bf Risk --}
The whole space of potential designs of a specification with respect to the
	$\abf$-theory.
\end{definition}
The definition of Risk leads to the risk matrix in
Figure~\ref{fig:quantitative}, defined as follows.
\begin{definition}{\bf Risk Matrix --}
	The risk matrix is a function between the three relations
	$s=\langle\rcc(\factRegion,\beliefRegion),\rcc(\factRegion,\assertionRegion),\rcc(\beliefRegion,\assertionRegion)\rangle$,
	where the maximum risk is defined by the $DR$ relation between the
	three groups of Regions, and the minimum risk by the $EQ$ relation over
	the same Regions. In between the two extremes, the granularity of
	possible intermediate configuration is defined by the calculus used
	(RCC5 in our case).
\end{definition}
While a risk matrix is often defined as a function between 
likelihood and impact of attacks, we suggest
that security weaknesses are all equally likely to be exploited
if there's a connection between the weakness and the asset that
an attacker wants to impact. Therefore, a risk matrix should
capture the number of insecure configurations of a system, rather
than predicating over likelihoods of weaknesses exploitation.

@@

\begin{theorem}{Port Weaknesses.}
From our definition of ports, it follows that there exist only the following \emph{six}
types of weaknesses, generating six types of insecure port in RCC5 (the
notation is reported for input-ports, i.e. on the left-hand side of the arrow):
\begin{enumerate}[start=1, label={W\arabic*)}]
	\item \emph{replace port}, where assertions reaches
		the port but do not pass the boundary of the agent (i.e. do not
		become belief of the agent)
	\item \emph{drop port}, where assertions reaches the
		port but do not pass the boundary of the agent (i.e. do not
		become belief of the agent)
	\item \emph{insertion port}, where some new
		information is believed by $a$ as incoming from the port but
		$s$ didn't send it
	\item \emph{injection port}, where information
		coming from $s$ is substituted with new information which
		becomes believed by $a$
	\item \emph{selective port}, where some information passes the port and
		part is either:
	\begin{enumerate}[start=1, label={W5.\arabic*)}]
		\item drop, 
		\item drop+insert.
	\end{enumerate}
\end{enumerate}
\end{theorem}

@@

\begin{enumerate}[start=6, label={W\arabic*)}]
	\item \emph{secure output port} and \emph{input drop port}
	\item \emph{secure output port} and \emph{input insertion port}
	\item \emph{output drop port} and \emph{input drop port}
	\item \emph{output drop port} and \emph{input insertion port}
	\item \emph{output drop port} and \emph{input secure port}
	\item \emph{output injection port} and \emph{input secure port}
\end{enumerate}

@@

\section*{Acknowledgment}
We thank Katia Santac\`a for the help in the development of the initial ideas, and for
the helpful discussions that allowed us to make this step forward.
